Learning Multiple Networks via Supervised Tensor Decomposition

### 2020-10-03 tensor_regress/NeurIPSw/neuripsw_tensor_v3.tex
v3

\section{Proposed models and motivating examples}
%For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
%\[ \sin \Theta(\mA,\mB)=\snormSize{}{\mA^T \mB^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB^\perp)\right\}. \]
Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor. Suppose the side information is available on each of the $K$ modes. Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$. We assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
\begin{align}\label{eq:decomp}
\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1, \ldots, \mX_K\mM_K\}\right),\notag\\
\text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K.
\end{align}
where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns, where $r_k\leq p_k$ for all $k\in[K]$, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$. Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.

@@

\begin{example}[Network population model] 
Network response model is recently developed in the context of neuroimanig analysis. The goal is to study the relationship between network-valued response and the individual covariates. Suppose we observe $n$ i.i.d.\ observations $\{(\mY_i, \mx_i): i=1,\ldots,n\}$, where $\mY_i\in\{0,1\}^{d\times d}$ is the brain connectivity network on the $i$-th individual, and $\mx_i\in\mathbb{R}^p$ is the individual covariate such as age, gender, cognition, etc. The network-response model~\citep{rabusseau2016low} has the form
\begin{equation}\label{eq:network}
\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n
\end{equation}
where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest. The model~\eqref{eq:network} is also a special case of our tensor-response model, with covariates on the last mode of the tensor.

>>>>

\section{Proposed models and motivating examples}
%For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
%\[ \sin \Theta(\mA,\mB)=\snormSize{}{\mA^T \mB^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB^\perp)\right\}. \]
Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor. Suppose the side information is available on each of the $K$ modes. Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$. We assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
\begin{align}\label{eq:decomp}
\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1, \ldots, \mX_K\mM_K\}\right),\notag\\
\text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K.
\end{align}
where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns with $r_k\leq p_k$ for all $k\in[K]$, and $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$. Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.

@@

\end{example}
\begin{example}[Network population model] 
Network response model is recently developed in the context of neuroimanig analysis. The goal is to study the relationship between network-valued response and the individual covariates. Suppose we observe $n$ i.i.d.\ observations $\{(\mY_i, \mx_i): i=1,\ldots,n\}$, where $\mY_i\in\{0,1\}^{d\times d}$ is the brain connectivity network on the $i$-th individual, and $\mx_i\in\mathbb{R}^p$ is the individual covariate such as age, gender, cognition, etc. The network-response model~\citep{rabusseau2016low} has the form
\begin{equation}\label{eq:network}
\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n
\end{equation}
where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest. The model~\eqref{eq:network} is also a special case of our tensor-response model, with covariates on the last mode of the tensor. 
\end{example}
